---
header-includes:
- \usepackage{amssymb, amsmath, amsthm}
- \usepackage{tabu}
- \newcommand{\E}{\mathbb{E}}
- \newcommand{\var}{{\rm Var}}
- \newcommand{\N}{\mathcal{N}}
output: pdf_document
---

\noindent \begin{tabu} to \textwidth {@{}X[4 l] @{}X[r]}
  \textbf{HW 4}           & \\ 
  \textbf{MFE 431: Data Analytics and Machine Learning}   & \\ 
  \textbf{Professor Lars Lochstoer}         & \\
  \textbf{Group 9}          &\\
  \textbf{Students: Yuhua Deng, Xiahao Wang, Nupur Solanki, Haoxuan Tong}
\end{tabu}


## Qn 1

a.

(i)

```{r}
rm(list=ls())

# we need the foreign package to import data in different format
suppressMessages(require(foreign))
suppressMessages(require(data.table))
suppressMessages(require(knitr))
suppressMessages(require(ggplot2))
suppressMessages(require(lfe))
suppressMessages(require(glmnet))

# Download data and set as data.table
StockRetAcct_raw <- as.data.table(read.dta("StockRetAcct_insample.dta"))

# Find excess return 
StockRetAcct_raw[,ExRet:=exp(lnAnnRet) - exp(lnRf)]

# Get the four variables
StockRetAcct <- StockRetAcct_raw[,.(FirmID, year, ExRet, lnIssue, lnProf, lnInv, lnME)]

# Get the square of variables
StockRetAcct <- StockRetAcct[, `:=`(lnIssue2 = lnIssue^2, lnProf2 = lnProf^2, lnInv2 = lnInv^2, lnME2 = lnME^2)]

# multiply by lnMe
StockRetAcct <- StockRetAcct[, `:=`(lnIssue_ME = lnIssue * lnME, lnProf_ME = lnProf * lnME, lnInv_ME = lnInv * lnME)]

# demean the data
StockRetAcct[, `:=`(lnIssue = lnIssue - mean(lnIssue,na.rm = TRUE), 
                    lnProf = lnProf - mean(lnProf,na.rm = TRUE),
                    lnInv = lnInv - mean(lnInv,na.rm = TRUE),
                    lnME = lnME - mean(lnME,na.rm = TRUE),
                    lnIssue2 = lnIssue2 - mean(lnIssue2,na.rm = TRUE),
                    lnProf2 = lnProf2 - mean(lnProf2,na.rm = TRUE),
                    lnInv2 = lnInv2 - mean(lnInv2,na.rm = TRUE),
                    lnME2 = lnME2 - mean(lnME2,na.rm = TRUE),
                    lnIssue_ME = lnIssue_ME - mean(lnIssue_ME,na.rm = TRUE),
                    lnProf_ME = lnProf_ME - mean(lnProf_ME,na.rm = TRUE),
                    lnInv_ME = lnInv_ME - mean(lnInv_ME,na.rm = TRUE)
                    ), by = year]

# add final column of 1's
StockRetAcct[, `:=`(Ones = 1)]

# sort by year
setkey(StockRetAcct, year)

# Run Machine learning based trading on page 46 of Lecture 4 
factor_return <- NULL
years <- sort(unique(StockRetAcct$year))

for(i in years){
  X <- StockRetAcct[year == i, c(15,4:14)]
  ex_ret <-  StockRetAcct[year == i, .(ExRet)]
  
  # replace NAs with 0 
  X[is.na(X)] <- 0
  # replace NAs with excess return
  ex_ret[is.na(ex_ret)] <- 0
  
  temp <- t(as.matrix(X)) %*% as.matrix(ex_ret)
  factor_return <- rbind(factor_return, t(temp))
}

# find the factor sample means
factor_ret_mean <- colMeans(factor_return)

# find the sample variance‐covariance matrix
sigma_cov <- cov(factor_return)

result <- list(Factor_mean = factor_ret_mean, var_cov = sigma_cov, SR_Return = colMeans(factor_return)/sqrt(diag(var(factor_return))))
result
```

(ii)
```{r}
rownames(factor_return) <- 1980:2014

crossValidation <- function(insample, outsample, alpha){
  
  insampleMean <- colMeans(insample)
  insampleCov <- cov(insample)
  fit <- glmnet(insampleCov,insampleMean, family = "gaussian", alpha = alpha)
  beta <- as.matrix(fit[['beta']])
  
  outsampleMean <- colMeans(outsample)
  outsampleCov <- cov(outsample)
  value <- outsampleCov %*% beta
  epsilon <- outsampleMean - value
  mse = apply(epsilon, 2, function(x) sum(x^2)/12)
  lambda <- fit$lambda
  return(list(mse=mse, beta=beta, lambda=lambda))
}

# find sample average factor returns and covariance matrix from 1980‐1999
get_coef_b <- function(pos, alpha){
  # 1st fold
  insample <- factor_return[(1+pos):(20+pos),]
  outsample <- factor_return[(21+pos):(25+pos),]
  fold_1 <- crossValidation(insample, outsample, alpha)
  
  # 2nd fold 
  insample <- factor_return[c((1+pos):(5+pos),(11+pos):(25+pos)),]
  outsample <- factor_return[(6+pos):(10+pos),]
  fold_2 <- crossValidation(insample, outsample, alpha)
  
  # 3rd fold 
  insample <- factor_return[c((1+pos):(10+pos), (16+pos):(25+pos)),]
  outsample <- factor_return[(11+pos):(15+pos),]
  fold_3 <- crossValidation(insample, outsample, alpha)
  
  # 4th fold
  insample <- factor_return[c((1+pos):(15+pos), (21+pos):(25+pos)),]
  outsample <- factor_return[(16+pos):(20+pos),]
  fold_4 <- crossValidation(insample, outsample, alpha)
  
  # fifth fold
  insample <- factor_return[(6+pos):(25+pos),]
  outsample <- factor_return[(1+pos):(5+pos),]
  fold_5 <- crossValidation(insample, outsample, alpha)
  
  lambdalist <- c(fold_1$lambda, fold_2$lambda, fold_3$lambda, fold_4$lambda, fold_5$lambda)
  mselist <- c(fold_1$mse, fold_2$mse, fold_3$mse, fold_4$mse, fold_5$mse)
  lambdaidx <- which.min(mselist)
  lambda_min <- lambdalist[lambdaidx]
  
  totalsample <- factor_return[(1+pos):(25+pos),]
  fitWholeSample <- glmnet(cov(totalsample),colMeans(totalsample), family = "gaussian", alpha = alpha, lambda = lambda_min)
  b <- fitWholeSample$beta
  return(b)
}

b <- get_coef_b(0, 0.5)

```

(iii)
```{r}
# find portfolio stats
outsample <- factor_return[26:35,]
MVE_ret <- as.matrix(t(b)) %*% t(outsample)
MVE_mean <- mean(MVE_ret)
MVE_SD <- sd(MVE_ret)
MVE_SR <- MVE_mean/MVE_SD
result_3 <- list(average_return = MVE_mean, standard_deviation= MVE_SD, sharpe_ratio = MVE_SR)
result_3
```

(iv)
```{r}
# get market return 
# Download data and set as data.table
stockdata <- as.data.table(read.dta("StockRetAcct_insample.dta"))

# Find excess return 
stockdata[,ExRet:=exp(lnAnnRet) - exp(lnRf)]

setkey(stockdata, year)

# value=weight
vwretd <- stockdata[, list(MeanExRetYr = weighted.mean(ExRet, MEwt, na.rm = T)), by = year]

# market std 
mktsd <- sd(vwretd$MeanExRetYr)
scale_facotor <- MVE_SD/mktsd
scale_MVE_ret <- MVE_ret/scale_facotor

vwretd_cum_ret <- cumprod(1 + vwretd$MeanExRetYr[26:35]) -1
MVE_cum_ret <- cumprod(scale_MVE_ret + 1) -1

plot(x =2005:2014, ylim=c(-0.3,2),y = vwretd_cum_ret, main="Return Estimation Vs Actual Return", type="l", col="blue", ylab="Return", xlab="Year")
lines(x =2005:2014 , y=MVE_cum_ret, type="l", col="red")
legend("topleft",legend=c("Estimation","Actual"),fill=c("red","blue"))
```

Now we could work on rolling window for prediction, 

```{r}
ret_list <- rep(0,10)
ret_list[1] <- MVE_ret[1]
alpha <- 0.5

for(i in 1:9){
  b <- get_coef_b(i, 0.5)
  outsample <- factor_return[26+i,]
  ret_list[(1+i)] <- outsample %*% as.matrix(b)
}

ret_sd <- sd(ret_list)
scale_facotor_3 <- ret_sd/mktsd
scale_MVE_ret_3 <- ret_list/scale_facotor_3

MVE_cum_ret_3 <- cumprod(scale_MVE_ret_3 + 1) -1

plot(x =2005:2014, ylim=c(-0.6,2),y = vwretd_cum_ret, main="Return Estimation Vs Actual Return With Rolling Window for One year Estimate", type="l", col="blue", ylab="Return", xlab="Year")
lines(x =2005:2014 , y=MVE_cum_ret_3, type="l", col="red")
legend("topleft",legend=c("Estimation","Actual"),fill=c("red","blue"))


```

Also we can manipulate the alpha to see the result for 10 year prediction

```{r}
library(RColorBrewer)
alpha <- seq(0,1,0.1)
ret_dif_alpha <- c()

for(i in alpha){
  b <- get_coef_b(0, i)
  outsample <- factor_return[26:35,]
  temp <- as.matrix(t(b)) %*% t(outsample)
  ret_dif_alpha <- rbind(ret_dif_alpha, temp)
}

df <- as.data.table(ret_dif_alpha)
stdDev <- apply(df, 1, sd)
df <- cbind(df, stdDev)

df[,scale_factor := stdDev/mktsd]

scale_factor <- df$scale_factor

normalized_ret <- apply(df, 2, function(x){ x/scale_factor } )

normalized_ret <- normalized_ret[, c(1:10)]

normalized_ret <- t(apply(normalized_ret, 1, function(x){ cumprod(x + 1) - 1}))

color <- brewer.pal(n = 10, name = 'Paired')

plot(x =2005:2014, ylim=c(-0.6,2),y = vwretd_cum_ret, main="Return Estimation Vs Actual Return 10 Year with different Alpha", type="l", col="black", ylab="Return", xlab="Year")
for(i in 1:10){
  lines(x =2005:2014 , y=normalized_ret[i,], type="l", col=color[i])  
}
legend("topleft", legend=paste0("alpha-",alpha), fill=color, cex = 0.8)
legend("bottomleft", legend="vwretd", fill=c("black"), cex = 0.8)


```

Also we can manipulate the alpha to see the result for 1 year prediction.
As, we can see, return is highest is when alpha = 0


```{r}
alpha <- seq(0,1,0.1)

one_yr_estimate <- function(alpha){
  ret_list <- rep(0,10)
  ret_list[1] <- MVE_ret[1]
  
  for(i in 1:9){
    b <- get_coef_b(i, alpha)
    outsample <- factor_return[26+i,]
    ret_list[(1+i)] <- outsample %*% as.matrix(b)
  }
  
  ret_sd <- sd(ret_list)
  scale_facotor_3 <- ret_sd/mktsd
  scale_MVE_ret_3 <- ret_list/scale_facotor_3
  
  MVE_cum_ret_3 <- cumprod(scale_MVE_ret_3 + 1) -1
  return(MVE_cum_ret_3)
}

cum_ret <- c()

for(i in alpha){
  temp <- one_yr_estimate(i)
  cum_ret <- rbind(cum_ret, temp)
}

colnames(cum_ret) <- years[26:35]

color <- brewer.pal(n = 10, name = 'Paired')

plot(x =2005:2014, ylim=c(-0.6,2),y = vwretd_cum_ret, main="Return Estimation Vs Actual Return 1 Year with different Alpha", type="l", col="black", ylab="Return", xlab="Year")
for(i in 1:10){
  lines(x =2005:2014 , y=cum_ret[i,], type="l", col=color[i])  
}
legend("topleft", legend=c(paste0("alpha-",alpha)), fill=color, cex = 0.8)
legend("bottomleft", legend="vwretd", fill=c("black"), cex = 0.8)

```

